# Reproducible Research: Peer Assessment 1 Submitted by Anandraj


## Loading and preprocessing the data
Load the data from file using read.csv()
```{r}
    df      <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day 
```{r}
stepsum <- aggregate(df$steps,list(df$date),sum) 
hist(stepsum$x,xlab = "Total Number of Step Taken per Day",main="Total Number of Steps per Day")
```
    
Mean and Median total number of steps taken per day 
Mean Steps taken 
```{r}
aggregate(x=df$steps,list(df$date),function(x) mean(x,na.rm=T)) 
```
Median of Steps taken
```{r}
aggregate(x=df$steps,list(df$date),function(x) median(x,na.rm=T))
```

## What is the average daily activity pattern?
Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
meanbyinterval <- aggregate(x=df$steps,list(df$interval),function(x) mean(x,na.rm=T)) 
plot(meanbyinterval$x ~ meanbyinterval$Group.1,type="l",xlab="Five minutes Interval",ylab="Average Number of Steps", main ="5-Min interval vs Average Steps")
```
    
    
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? 
```{r}
meanbyinterval[meanbyinterval$x==max(meanbyinterval$x),]$Group.1
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
    
To create the new dataset by imputing missing values copy original dataset df to dfout.
```{r}
dfout <- df 
```
To impute the values use a for loop to select the rows with missing values then assign the intervalmean to that value. 
```{r}
for (i in which(is.na(dfout$steps)))
{ 
dfout$steps[i] <- meanbyinterval[meanbyinterval$Group.1 == dfout[i,3],2] 
}
```
Histogram of the total number of steps taken each day after imputing values
```{r}
stepsum <- aggregate(dfout$steps,list(dfout$date),sum) 
hist(stepsum$x,xlab = "Total Number of Step Taken per Day",main="Total Number of Steps per Day")
```
Calculate and report the mean and median total number of steps taken per day 
Mean of Steps for each day
```{r}
aggregate(x=dfout$steps,list(dfout$date),function(x) mean(x,na.rm=T)) 
```
Median of Steps for each day
```{r}
aggregate(x=dfout$steps,list(dfout$date),function(x) median(x,na.rm=T))
```
     

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable to indicate if Weekend or Weekday
```{r}
wd <- rep("day",length(dfout$date))
wd[which(!(weekdays(as.Date(dfout$date,format="%Y-%m-%d")) %in% c("Saturday","Sunday")))] <- "Weekday"
wd[which(weekdays(as.Date(dfout$date,format="%Y-%m-%d")) %in% c("Saturday","Sunday"))] <- "Weekend"
wd <- as.factor(wd)
```
Average steps across 5-Min interval split by Weekday & Weekdends
```{r}
meanbyintday <- aggregate(x=dfout$steps,list(df$interval,wd),function(x) mean(x,na.rm=T))
```
Plot the difference in Pattern
```{r}
library("ggplot2")
qplot(Group.1,x,data=meanbyintday,geom="line",facets=Group.2~.,color=Group.2,xlab="5-Min Interval",ylab="Average Steps",main="Weekend vs Weekdays")
```